\section*{Abstract}
\label{sec:abstract}


\textbf{Motivation:}~Growth rate is a major component of the
evolutionary fitness of microbial organisms; when nutrients are
plentiful, faster-growing strains come to dominate populations,
whereas slower-growing strains are wiped out. Maintaining a high
growth rate is, therefore, a priority for cell lineages, making growth
rate an excellent surrogate for cell health. When measured in
different genetic backgrounds or environments, cell health can inform
about genetic interaction and drug sensitivity. In the high-throughput
procedures QFA and SGA, microbial cultures are inoculated in an array
on solid agar, and growth is measured to obtain quantitative estimates
of culture fitness. Neighbouring cultures, which are often different
genetic strains, grow at different rates, and consume nutrients at
different rates, creating gradients in nutrient density between
cultures. We believe that diffusion of nutrients between cultures
affects growth. Current analysis does not account for this; instead,
it is assumed that cultures grow independently. I use a network model
of nutrient diffusion and nutrient-dependent growth, to correct for
competition, to try to improve the accuracy and precision of growth
and fitness estimates. I test the model against QFA data from
experiments on \textit{Saccharomyces cerevisiae}. Ultimately, the
model might be used to estimate genetic interaction and drug
sensitivity with higher accuracy and precision.
\\
%Original
% \textbf{Motivation:}~Growth rate is a major component of the
% evolutionary fitness of microbial organisms.~When nutrients are
% plentiful, fast-growing strains come to dominate populations whereas
% slower-growing strains are wiped out. Maintaining growth rate is
% therefore a priority for cell lineages and this makes growth rate an
% excellent surrogate for the health of cells. Measuring the health of
% cells grown in different genetic backgrounds or environments can
% inform about genetic interaction and drug sensitivity. In
% high-throughput procedures such as QFA and SGA, arrays of microbial
% cultures are grown on solid agar plates and quantitative fitness
% estimates are determined from growth measurements. Diffusion of
% nutrients along gradients in nutrient density arising between fast-
% and slow-growing neighbours is likely to affect growth rate and
% fitness estimates. However, current analyses assume that cultures grow
% independently. I study data from QFA experiments growing
% \textit{Saccharomyces cerevisiae} to test a mass action kinetic model
% of nutrient dependent growth and diffusion across a network. I try to
% correct for competition to provide more accurate and precise fitness
% estimates to allow more accurate and precise genetic interaction and
% drug screens.
% \\
\textbf{Results:}~The competition model fits timecourses from a
384-format QFA plate inoculated with dilute cultures, with similar
closeness to a current QFA model. It does so using far fewer
parameters (387 vs 1152). Closeness of fit is improved for fast
growing strains. Fitness rankings of strains on the same plate agree
with a previous QFA study and independent experiments in the positions
of the fastest and slowest growing strains, but disagree for medium
growing strains. Estimates are less precise for the fastest growing
strains, but more precise for 36 out of 50 strains. In-cross plate
calibration and validation, using QFA plates with a higher inoculum
density, the model overcorrects for competition effects to a similar
degree that current models undercorrect. A different method of fitting
is required to find globally optimal solutions.
\\

\textbf{Results:}~Using far fewer parameters (387 vs 1152), the
proposed model fits timecourses from a 384-format QFA plate inoculated
with dilute cultures with a similar accuracy to a current QFA
model. Accuracy is improved for fast growing strains. Fitness rankings
of strains on the same plate agree with a previous QFA study and
independent experiments in the positions of the fastest and slowest
growing strains, but disagree for medium growing strains. Estimates
are less presise for the fastest growing strains, but more precise for
36 out of 50 strains. In cross plate callibration and validation,
using QFA plates with a higher inoculum density, the model
overcorrects for competition effects to a similar degree that current
models undercorrect. A different method of fitting is required to find
globally optimal solutions.
\\
\textbf{Availability and Implementation:}~CANS, a Python package
developed for the analysis in this paper, is freely available at
\href{https://github.com/lwlss/CANS}{https://github.com/lwlss/CANS}.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "report"
%%% End:
